Towards a Machine Learning approach for Aggregate Computing

### 2021-11-04 paper-2021-woa-ac-ml.tex
refactor the abstract, pass on intro

\maketitle
\begin{abstract}
\suggestions{}
Artificial intelligence is here to stay. 
 In particular, Deep Learning techniques in the last years outperform 
 hand-craft human solutions in different contexts, ranging from computer vision to videogames.
%
Currently, though, an all-encompassing Deep learning algorithm does not exist, 
 so researchers explore the possibility to apply these techniques in different fields.

One of the most challenging is Collective-Adaptive System (CAS). % Or do we want to use Cyber-Physical Swarm??
 Here, engineers have to handle large scale, global-to-local behaviour mapping,
 highly environment stochasticity. 
%
Historically, researchers favour handling them with self-organisation,
 by which a collective behaviour emerges from continuously local interaction between simple components.
%
A novel technique to specify self-organising behaviour in a composable and functional manner is Aggregate Computing. 
 Even though it is promising for expressing high-level collective behaviour, 
 it needs to build low-level blocks that tend to be tricky.

On the opposite side, some research lines apply Deep learning (supervised, reinforcement, non-supervised and self-supervised)
 along with CAS. Currently, though, no one gain popularity and solution tend to be application-specific.
%
So, we outline a Machine Learning technique that could be applied in Aggregate Computing, 
 making the possibility to combine specification-based and automatic approaches that can easily scale in application complexity and to any system size.
\end{abstract}
\begin{IEEEkeywords}
  Collective-Adaptive System,
  Aggregate Computing,
  Deep Learning,
  Reinforcement Learning
\end{IEEEkeywords}

@@

CASs are characterised by 
\begin{itemize}
\item a global order that emerges from local node interactions, 
\item  decentralised control, and 
\item adaptive behaviours towards environmental changes to pursuit a collective goal.
\end{itemize}
CAS are notoriously hard to design due as part of complex systems. 
%
So, over the years, in literature, various approaches aim at handling this complexity. 
%
Some of them take inspiration from the natural system
 composed of social animals --- like bees, fish and ants. 
%
Here, a sort of hive mind arises -- 
 something referred to as swarm intelligence -- 
born from local interactions between agents. 
%
This kind of behaviour is called self-organisation, 
 namely a global order that emerges from continuous 
 interactions between simple entities.

@@

In this direction, Aggregate Computing is an innovative solution by which 
 developers can express collective self-organising behaviours through a functional program specification.
%
Indeed, an aggregate blueprint consists of the manipulation of \textit{computational fields}, a distributed
 data structure where each point in the spacetime is associated with the result of a computation
 performed by a node.
%
These operations are defined in \textit{Field-calculus} --- the root of Aggregate Computing. 
 This calculus describes a set of minimal constructs to express any spatiotemporal computation.
%
Thanks to the ``composable'' nature of Aggregate Computing, we find common abstractions 
 in collective behaviour specification, and then we encapsulate them in so-called ``building blocks''.
%
With those, different properties have been proven (such as \textit{eventual consistency}) and various
 applications have been built (such as swarm robotics, crowds engineering).

@@

So, we claim that Machine Learning can help us in handling this "unseen" situation, leaving only the burden
 to developers of defining the right collective specifications.
%
This lead to a "hybrid" approach where a part of behaviour is still expressed using common 
 Aggregate Computing abstractions and another part is distilled through learning.
%
In this paper, as the first steps towards \hybridaggregate{}, we investigate how Aggregate Computing can be
 extended with Reinforcement Learning (and in particular with Independent Reinforcement Learning) through a case
 study where we outline the performance improvement of the "hop count" program.
\suggestions{If we leave the part about "roadmap" and "machine learning similar setting", we need to add some description here.}
\suggestions{I do not know if we can avoid it altogether, but also a brief description of machine learning approaches can be useful in the paper flow}
\todos{Outline}

@@

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{img/1}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{img/50}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{img/100}
  \end{subfigure}
  \caption{The error evolution compared with naive implementation. The \textit{black} line 
  describes the Reinforcement Learning solution error. The \textit{yellow} line shows
  the naive solution error.}
  \label{fig:simulation}
\end{figure}
\todos{add blinking scenarios}
\todos{add movement scenarios}

@@

\bibliographystyle{IEEEtran}
\bibliography{biblio}
%%
%% If your work has an appendix, this is the place to put it.
\appendix

>>>>

\maketitle
\begin{abstract}
Artificial intelligence is here to stay. 
 In particular, Machine Learning techniques in the last years outperform 
 hand-craft human solutions in different contexts, ranging from computer vision to videogames.
%
Currently, though, an all-encompassing Machine learning algorithm does not exist, 
 so researchers explore the possibility to apply these techniques in different fields.

One of the most challenging is Collective-Adaptive System (CAS). % Or do we want to use Cyber-Physical Swarm??
 Here, engineers have to handle large scale, global-to-local behaviour mapping,
 highly environment stochasticity. 
%
Historically, researchers favour handling them with self-organisation,
 by which a collective behaviour emerges from continuously local interaction between simple components.
%
A novel technique to specify self-organising behaviour in a composable and functional manner is Aggregate Computing.
 In particular, this paradigm enables collective program specification as a whole, 
 as it is based on a distributed data structure manipulation called \textit{computational fields}.
 It is successfully used in different scenarios ranging from Swarm Robotics to Smart cities.

Even though it is promising for expressing high-level collective behaviour,
 practically it is difficult to make programs work in highly dynamic networks, 
 where nodes have to deal with continuous changes at the environment side.
%
Therefore, with this work, we stand the first step towards integration between Aggregate Computing and Machine Learning,
 in order to improve adaptivity acting in the low-level specifications but maintaining the high-level program behaviours.

In the article, we outline issues, opportunities and methodologies of these integrations, presenting a case study
 where Aggregate Computing is successfully combined with Hysteretic Q-Learning.
%

\end{abstract}
\begin{IEEEkeywords}
  Collective-Adaptive System,
  Aggregate Computing,
  Deep Learning,
  Reinforcement Learning
\end{IEEEkeywords}

@@

CAS are notoriously hard~\cite{DBLP:conf/huc/Ferscha15} to design as they are characterised by 
\begin{itemize}
\item a global order that emerges from local node interactions, 
\item decentralised control, and 
\item adaptive behaviours towards environmental changes to pursuit a collective goal.
\end{itemize} 
%
So, over the years, in literature, various approaches aim at handling this complexity. 
%
Some of them take inspiration from the natural system
 composed of social animals --- like bees, fish and ants. 
%
Here, a sort of hive mind arises -- 
 something referred to as swarm intelligence -- 
born from local interactions between agents. 
%
These kinds of behaviours are accomplished with self-organisation, 
 namely a global order that emerges from continuous 
 interactions between simple entities.

@@

In this direction, Aggregate Computing~\cite{DBLP:journals/computer/BealPV15} is an innovative solution by which 
 developers can express collective self-organising behaviours through a functional program specification.
%
Indeed, an aggregate program consists of the manipulation of \textit{computational fields}, a distributed
 data structure where each point in the spacetime is associated with the result of a computation
 performed by a node.
%
These operations are defined in \textit{Field-calculus}~\cite{DBLP:conf/coordination/AudritoBDV18} --- the root of Aggregate Computing. 
 This calculus describes a set of minimal constructs to express any spatiotemporal computation.
%
Thanks to the ``composable'' nature of Aggregate Computing, we find common abstractions 
 in collective behaviour specification, and then we encapsulate them in so-called ``building blocks''.
%
With those, different properties have been proven (such as \textit{eventual consistency}~\cite{DBLP:journals/taas/BealVPD17}) and various
 applications have been built (such as swarm robotics, crowds engineering~\cite{DBLP:journals/eaai/CasadeiVAPD21}).

@@

So, we claim that Machine Learning can help us in handling this "unseen" situation, leaving only the burden
 to developers of defining the right collective specifications.
%
This lead to an ``hybrid" approach where a part of behaviour is still expressed using common 
 Aggregate Computing abstractions and another part is distilled through learning~\cite{research}.
%
In this paper, as the first steps towards a so-called \hybridaggregate{}, we investigate how Aggregate Computing can be
 extended with Reinforcement Learning through a case
 study where we outline the performance improvement of the "hop count" program.
\suggestions{If we leave the part about "roadmap" and "machine learning similar setting", we need to add some description here.}
\suggestions{I do not know if we can avoid it altogether, but also a brief description of machine learning approaches can be useful in the paper flow}

@@

\begin{figure*}
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{img/1}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{img/50}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{img/100}
  \end{subfigure}
  \caption{The error evolution compared with naive implementation. The \textit{black} line 
  describes the Reinforcement Learning solution error. The \textit{yellow} line shows
  the naive solution error.}
  \label{fig:simulation}
\end{figure*}
%\todos{add blinking scenarios}
%\todos{add movement scenarios}

@@

\bibliographystyle{IEEEtran}
\bibliography{biblio}
%%
%% If your work has an appendix, this is the place to put it.
%\appendix
